Differential invariant signatures for planar Lie group transformations with application to images

### 2022-03-09 paper/newerpaper.tex
And one more..
revisions for referee

% Load whatever packages you need
\usepackage{amsmath,amsfonts,graphicx}
\def\R{\mathbb{R}}
\usepackage{mathtools}
\usepackage{xcolor}
\usepackage{todonotes}
\usepackage{tikz}
\usepackage{array}
\usepackage{natbib}

@@

\begin{document}
\maketitle
\begin{abstract}
The actions of different various Lie groups underlie the change of appearance of objects in images as the viewpoint changes, e.g., through camera motion. Despite significant advances in object recognition using machine learning in recent years, the question of how to recognise an object in an image as its appearance varies through camera motion and similar effects remains open. We demonstrate how differential invariant signatures can be derived for each of the transformation groups, and how the underlying invariances reflect the group-subgroup structure of the relevant Lie groups.

@@

In order to define differential invariants, it is necessary to identify how the group transformation $\varphi$ acts on derivatives of the function $f$ via the jet space~\citep{Kruglikov2011,OlverEIS}. The $d$-th order jet space $J^d$ consists of the function itself and the set of all partial derivatives of the function up to order $d$. (Note that there are $\left( \begin{array}{c} d+k-1 \\ k \end{array} \right)$ partial derivatives of order $d$.) As the group $\mathcal{G}$ is local, it acts naturally on derivatives of the transformation, and so each individual transformation $\varphi \in \mathcal{G}$ can be prolonged to act on the jet space $J^d$. We can thus prolong the action of the group and consider the action of $\mathcal{G}$ on not just $f$, but all derivatives of $f$ up to order $d$. A $d$-th order differential invariant is then a local scalar function $I: J^{d} (f,\Omega, k) \to \mathbb{R}$ that is invariant under the action of $\mathcal{G}$, i.e., $I(\varphi(f,\Omega, k)) = I(f, \Omega, k) \forall \varphi \in \mathcal{G}$. It can also be useful to consider the maximum order of derivative in the invariant, which we term degree, e.g., $f_{xy}^2 f_x f_y$ has degree 2.

@@

The group that has been of particular interest for image analysis is the projective group, because of its ubiquity in camera models. Invariants for this group have been computed using points~\citep{Gros1992,Suk2000}, geometric moments~\citep{Suk2004}, curvature~\citep{Hann2002a}, and differential and joint invariants~\citep{Hubert2007,Arora2009,Kogan2014,Li19}.

Although there are computational challenges regarding the numerical approximation of derivatives on images, the fact that differential invariants are robust to occlusion, cheap to compute, and well-behaved with respect to small changes in the image, has made them popular for researchers in image analysis~\citep{Florack1993}. Two particular tools have been key to this: Cartan's moving frame~\citep{Cartan35} (see Section~\ref{sec:movingframes}), which was used to construct image invariants as early as 1994 by \citet{Faugeras1994}, and the concept of an image signature built from a set of invariants~\citep{Calabi1998}. An algorithm to compute differential and joint invariants and determine complete signature sets is described by \citet{Olver2005,Olver2007}
The entire set of invariants is computed using the method of moving frames and then the signature set of a maximal set of functionally independent invariants is identified. A third step is possible, in which the dimension of the signature set is reduced by eliminating those invariants whose values can be computed from the values of derivatives of some subset of the invariants. Further work, including applications to ODEs and the Calculus of Variations can be found in~\cite{Mansfield}.

@@

The function is depicted as a greyscale image in Fig.~\ref{fig:function}.
We will use this function throughout, using the grid on the right of the figure to help visualise both the transformation and how the image
is mapped into the three-dimensional signature surface.

@@

A better approach is to take three invariants of the same weight.
Taking $I_1^2$, $I_2^2$, and $I_3$ from \eqref{eq:e2tensor}, all of which are
weight $4$, they can be projected on to the unit sphere, which can be
multiplied by $I_0 = f$ to form a signature surface. This signature is
defined by:
\begin{equation}
    \label{eq:sim2signature}
\mathcal{I}_{\text{Sim(2)}}(f) = \frac{f}{\sqrt{I_1^4 + I_2^4 + I_3^2}}(I_1^2, I_2^2,
        I_3) .
\end{equation}
This signature is only singular when $f_x = f_y = f_{xx} = f_{yy} = f_{xy}
= 0$. A sample similarity transformation and computed signature is shown in
Fig.~\ref{fig:Sim2}.

@@

A moving frame is a mapping into a group $\mathcal{G}$ that provides a frame of reference along a manifold.  It was extended by~\citet{Cartan35} to the `method of moving frames'. The actions of a Lie group $\mathcal{G}$ enables the identification of invariants to that group through the definition of a $\mathcal{G}$-equivariant mapping. A (right-) moving frame is a smooth $\mathcal{G}$-equivariant map $\rho : \mathcal{M} \to \mathcal{G}$ such that $\rho (g \cdot z) = \rho(z) \cdot g^{-1}$, where $z \in J^d$ and where $\cdot$ denotes the group operation of $\mathcal{G}$. It is necessary~\citep{Olver2013} for the group to act freely. This means that the group action needs to be extended to the entire jet space of derivatives up to order $d$ by implicit differentiation of the group action, a process known as prolongation.

@@

These invariants are a functionally independent, complete, set of
polynomial differential invariants up to second derivatives. Note that the
invariants found previously by tensor contraction Eq. (\ref{eq:e2tensor}) in
Section~\ref{sec:contraction} can be expressed in terms of them:
\begin{align*}
  f = f &= K_0 \\
  f_if_i = f_x^2 + f_y^2 &= K_1 \\
  f_{ii} = f_{xx} + f_{yy} &= \frac{K_2 + K_4}{K_1} \\
  f_{ij} f_{ij} = f_{xx}^2 + 2f_{xy}^2 + f_{yy}^2 &= \frac{K_2^2 + K_4^2 + 2K_3^2}{K_1^2} \\
  f_i f_j f_{ij} = f_x^2 f_{xx} + 2f_x f_y f_{xy} + f_y^2 f_{yy} &= K_2.
\end{align*}

@@

The moving frame calculation is then:
\begin{enumerate}
\item The frame $\bar f_{\bar x}=1$, $\bar f_{\bar y}=0$ determines two group parameters:
$$ a = f_x/(f_x^2 + f_y^2),\quad b = f_y/(f_x^2 f_y^2).$$

\item The frame $\bar f_{\bar x \bar y}=\bar f_{\bar y \bar y}=0$ determines the next group parameters:
$$ c = (2 f_xf_yf_{xyy} - f_{xx}f_y^2 - f_{yy}f_x^2)/(2 f_x^2 + f_y^2)^2$$
$$ d = (f_{xy}f_x^2 - f_x f_y f_{xx} - f_{xy}f_y^2 - f_{yy} f_x f_y)/(2 f_x^2 + f_y^2)^2.$$

@@

We begin the construction of the moving frame by choosing $\bar f_{\bar x}
= 1$ and $\bar f_{\bar y}=  0$, which determines the group parameters $a =
(1-c f_y)/f_x.$ and $b = -d f_y/f_x$. Direction substition then shows that
$\bar f_{\bar y \bar y} = D d^2 / f_x^2$, where $D = f_{x}^2 f_{yy} + f_y^2
f_{xx} - 2 f_x f_y f_{xyy}$ was listed in Eq.~\eqref{eq:cde} as an invariant of
$A(2)$, which is a subgroup of $PSL(3,\R)$. Note that $D$ and
$\bar f_{\bar y \bar y}$ have the same sign. If $D < 0$, we choose the
frame $\bar{f}_{\bar y \bar y} = -1$ and if $D \ge 0$ we choose $\bar
f_{\bar y \bar y} = 1$. Then $d = f_x/|D|^{1/2}.$

The frame $\bar f_{ \bar y \bar y \bar y} = 0$ determines the group
parameter $ h = \frac{f_{yyy}f_x^3 - 3 f_x^2 f_y f_{xyy} + 3 f_x f_y^2
f_{xxy} - f_y^3 f_{xxx}}{6|D|^{3/2}},$ while $\bar f_{ \bar x \bar y 
y}=0$ provides $ c = \frac{h f_x^2}{d D} - \frac{f_x f_{xy} - f_y
f_{xx}}{D}$, and $\bar f_{ \bar x \bar x}=0$ gives $ g = \frac{1}{2
f_x^2}(f_{xx} + 2 c ( f_x f_{xy} - f_y f_{xx} ) + c^2 D).$ In fact, the
numerator of the expression for $h$ was another invariant of $A(2)$ labelled as
$E$ in Eq.~\eqref{eq:cde}.

>>>>

% Load whatever packages you need
\usepackage{amsmath,amsfonts,graphicx}
\def\R{\mathbb{R}}
\usepackage{mathtools}
\usepackage{xcolor}
\usepackage{todonotes}
\usepackage{tikz}
\usepackage{array}
\usepackage{natbib}
\usepackage{url}

@@

\begin{document}
\maketitle
\begin{abstract}
The actions of various Lie groups underlie the change of appearance of objects in images as the viewpoint changes, e.g., through camera motion. Despite significant advances in object recognition using machine learning in recent years, the question of how to recognise an object in an image as its appearance varies through camera motion and similar effects remains open. We demonstrate how differential invariant signatures can be derived for each of the transformation groups, and how the underlying invariances reflect the group-subgroup structure of the relevant Lie groups.

@@

In order to define differential invariants, it is necessary to identify how the group transformation $\varphi$ acts on derivatives of the function $f$ via the jet space~\citep{Kruglikov2011,OlverEIS}. The $d$-th order jet space $J^d$ is coordinatised by the function itself and the set of all partial derivatives of the function up to order $d$. (Note that there are $\left( \begin{array}{c} d+k-1 \\ k \end{array} \right)$ partial derivatives of order $d$.) As the group $\mathcal{G}$ is local, it acts naturally on derivatives of the transformation, and so each individual transformation $\varphi \in \mathcal{G}$ can be prolonged to act on the jet space $J^d$. We can thus prolong the action of the group and consider the action of $\mathcal{G}$ on not just $f$, but all derivatives of $f$ up to order $d$. A $d$-th order differential invariant is then a local scalar function $I: J^{d} (f,\Omega, k) \to \mathbb{R}$ that is invariant under the action of $\mathcal{G}$, i.e., $I(\varphi(f,\Omega, k)) = I(f, \Omega, k) \forall \varphi \in \mathcal{G}$. It can also be useful to consider the maximum order of derivative in the invariant, which we term the degree, e.g., $f_{xy}^2 f_x f_y$ has degree 2.

@@

The group that has been of particular interest for image analysis is the
projective group, because of its ubiquity in camera models. Invariants for
this group have been computed using points~\citep{Gros1992,Suk2000},
geometric moments~\citep{Suk2004}, curvature~\citep{Hann2002a}, and
differential and joint
invariants~\citep{Hubert2007,Arora2009,Kogan2014,Li19}.  In particular
Olver, in \citet{Olver2020}, uses the same approach as we do in
\S\ref{sec:projective}, but with a different moving frame normalization to
produce a different set of projective image invariants.

Although there are computational challenges regarding the numerical approximation of derivatives on images, the fact that differential invariants are robust to occlusion, cheap to compute, and are based on local computations, has made them popular for researchers in image analysis~\citep{Florack1993}. Two particular tools have been key to this: Cartan's moving frame~\citep{Cartan35} (see Section~\ref{sec:movingframes}), which was used to construct image invariants as early as 1994 by \citet{Faugeras1994}, and the concept of an image signature built from a set of invariants~\citep{Calabi1998}. An algorithm to compute differential and joint invariants and determine complete signature sets is described by \citet{Olver2005,Olver2007}
The entire set of invariants is computed using the method of moving frames and then the signature set of a maximal set of functionally independent invariants is identified. A third step is possible, in which the dimension of the signature set is reduced by eliminating those invariants whose values can be computed from the values of derivatives of some subset of the invariants. Further work, including applications to ODEs and the Calculus of Variations can be found in~\cite{Mansfield}.

@@

The function is depicted as a greyscale image in Fig.~\ref{fig:function}.
We will use this function throughout, using the grid on the right of the figure to help visualise both the transformation and how the image
is mapped into the three-dimensional signature surface. Note that there is nothing particularly special about this particular function; it has relatively 
well-behaved derivatives, includes a critical point, and decays towards the boundary of the image.

@@

A better approach is to take three invariants of the same weight.
Taking $I_1^2$, $I_2^2$, and $I_3$ from \eqref{eq:e2tensor}, all of which are
weight $4$, defines a surface up to scale, and so they can be projected on to the unit sphere. 
We therefore multiply by another invariant to form the signature surface; an appropriate choice 
is $I_0 = f$. This signature is defined by:
\begin{equation}
    \label{eq:sim2signature}
\mathcal{I}_{\text{Sim(2)}}(f) = \frac{f}{\sqrt{I_1^4 + I_2^4 + I_3^2}}(I_1^2, I_2^2,
        I_3) .
\end{equation}
It is only singular when $f_x = f_y = f_{xx} = f_{yy} = f_{xy}
= 0$. A sample similarity transformation and computed signature is shown in
Fig.~\ref{fig:Sim2}.

@@

A moving frame is a mapping into a group $\mathcal{G}$ that provides a frame of reference along a manifold. It was used by~\citet{Cartan35} as the basis for his `method of moving frames', which uses Cartan equivalence to identify the set of structural invariants of a manifold. \cite{Fels1998,Fels1999} showed that the actions of a Lie group $\mathcal{G}$ enable the identification of invariants to that group through the definition of a $\mathcal{G}$-equivariant mapping. A (right-) moving frame is a smooth $\mathcal{G}$-equivariant map $\rho : \mathcal{M} \to \mathcal{G}$ such that $\rho (g \cdot z) = \rho(z) \cdot g^{-1}$, where $z \in J^d$ and where $\cdot$ denotes the group operation of $\mathcal{G}$. It is necessary~\citep{Olver2013} for the group to act freely. This means that the group action needs to be extended to the entire jet space of derivatives up to order $d$ by implicit differentiation of the group action, a process known as prolongation.

@@

These invariants are a functionally independent, complete, set of
polynomial differential invariants up to second derivatives. Note that the
invariants found previously by tensor contraction Eq. (\ref{eq:e2tensor}) in
Section~\ref{sec:contraction} can be expressed in terms of them (see the replacement theorem in \cite{Fels1999} for a justification of this):
\begin{align*}
  f = f &= K_0 \\
  f_if_i = f_x^2 + f_y^2 &= K_1 \\
  f_{ii} = f_{xx} + f_{yy} &= \frac{K_2 + K_4}{K_1} \\
  f_{ij} f_{ij} = f_{xx}^2 + 2f_{xy}^2 + f_{yy}^2 &= \frac{K_2^2 + K_4^2 + 2K_3^2}{K_1^2} \\
  f_i f_j f_{ij} = f_x^2 f_{xx} + 2f_x f_y f_{xy} + f_y^2 f_{yy} &= K_2.
\end{align*}

@@

The moving frame calculation is then:
\begin{enumerate}
\item The partial cross-section $\bar f_{\bar x}=1$, $\bar f_{\bar y}=0$ determines two group parameters:
$$ a = f_x/(f_x^2 + f_y^2),\quad b = f_y/(f_x^2 f_y^2).$$

\item The partial cross-section $\bar f_{\bar x \bar y}=\bar f_{\bar y \bar y}=0$ determines the next group parameters:
$$ c = (2 f_xf_yf_{xyy} - f_{xx}f_y^2 - f_{yy}f_x^2)/(2 f_x^2 + f_y^2)^2$$
$$ d = (f_{xy}f_x^2 - f_x f_y f_{xx} - f_{xy}f_y^2 - f_{yy} f_x f_y)/(2 f_x^2 + f_y^2)^2.$$

@@

We begin the construction of the moving frame by choosing $\bar f_{\bar x}
= 1$ and $\bar f_{\bar y}=  0$, which determines the group parameters $a =
(1-c f_y)/f_x.$ and $b = -d f_y/f_x$. Direct substition then shows that
$\bar f_{\bar y \bar y} = D d^2 / f_x^2$, where $D = f_{x}^2 f_{yy} + f_y^2
f_{xx} - 2 f_x f_y f_{xyy}$ was listed in Eq.~\eqref{eq:cde} as an invariant of
$A(2)$, which is a subgroup of $PSL(3,\R)$. Note that $D$ and
$\bar f_{\bar y \bar y}$ have the same sign. If $D < 0$, we choose the
frame $\bar{f}_{\bar y \bar y} = -1$ and if $D \ge 0$ we choose $\bar
f_{\bar y \bar y} = 1$. Then $d = f_x/|D|^{1/2}.$

The partial cross-section $\bar f_{ \bar y \bar y \bar y} = 0$ determines the group
parameter $ h = \frac{f_{yyy}f_x^3 - 3 f_x^2 f_y f_{xyy} + 3 f_x f_y^2
f_{xxy} - f_y^3 f_{xxx}}{6|D|^{3/2}},$ while $\bar f_{ \bar x \bar y 
y}=0$ provides $ c = \frac{h f_x^2}{d D} - \frac{f_x f_{xy} - f_y
f_{xx}}{D}$, and $\bar f_{ \bar x \bar x}=0$ gives $ g = \frac{1}{2
f_x^2}(f_{xx} + 2 c ( f_x f_{xy} - f_y f_{xx} ) + c^2 D).$ In fact, the
numerator of the expression for $h$ was another invariant of $A(2)$ labelled as
$E$ in Eq.~\eqref{eq:cde}.
